---
title: Analysis of USAF base locations, state populations, and UFO sightings
author: Mackenzie Gray
---

# Loading and cleaning

First import all of the necessary packages

```{r, message=F, warning=F}
library(lmtest)
library(plyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(GGally)
library(lubridate)
library(zoo)
```

Next, read in the necessary data.
`sightings` is the ufo sighting records from NUFORC, `base_locs` location information about USAF bases in the US,
`name_to_abbr` a mapping of US state names to abbreviations, and `stat_pops` US Census population data per state per decade from 1960 to 2010. 

```{r, message=F, warning=F}
sightings <- read_csv("../../data/raw/ufo_sightings.csv")
base_locs <- read_csv("../../data/raw/usaf_base_locs.csv")
state_pops <- read_csv("../../data/raw/state_pops.csv")
```

Next, transform and select the relevant data.

```{r}
sightings <- sightings  %>%
    filter(X1 < 109088) %>%
    filter(state %in% state.abb)

state_pops$Name <- mapvalues(state_pops$Name, state.name, state.abb)
state_pops <- state_pops %>%
    filter(Name %in% state.abb) %>%
    select(one_of(c("1970", "1980", "1990", "2000", "2010"))) %>%
    apply(1, mean)

sightings <- sightings  %>%
    group_by(state) %>%
    count()

base_locs <- base_locs  %>%
    group_by(State) %>%
    count() %>%
    merge(state.abb, by.x="State", by.y=1, all=T) %>%
    mutate(n=replace(n, is.na(n), 0)) %>%
    filter(!is.na(State))

df <- data.frame(sightings=sightings$n, base_count=base_locs$n, pop=state_pops, name=state.abb)

```

## Statistical analysis
Now that the data is all clean and together, we can begin analysis. 

### Initial EDA

We can create a pairplot of the quantitative variables to get a rough look at how population and number of Air Force bases relate to the number of UFO sightings.
It is reasonable to assume that states with a higher population will have more bases and more sightings just by virtue of having more people living there.
This can be confirmed using `ggpairs` from the `GGally` package.

```{r}
ggpairs(df[,c("base_count", "pop", "sightings")])
```

From this we can make a few initial observations.
It seems as though `pop` is relatively correlated with both `sightings` and `base_count` ($R^2 = 0.456$ and $0.338$, respectively).
The relationship between `base_count` and `pop` does not appear entirely linear however.
Additionally, it seems as though this relationship may have some degree of hetereoscedasticity.
It also appears, however, that `base_count` and `sightings` are correlated with $R^2 = 0.315$.
This suggests multicolinearlity between the predictor variables `base_count` and `pop`, which will complicate matters further on.

### State populations vs number of sightings

Although these two variables seems obviously related, it is good practice to double check assumptions.

```{r}
(s1 <- summary(fit1 <- lm(sightings ~ pop, data=df)))
```

Given $R^2 = `r round(s1$r.squared, 3)`$, $\beta_{pop} = `r round(s1$coefficients[[2, 1]], 5)`$, and a 95% confidence interval for $\beta_{pop} = (`r round(confint(fit1)[2,1], 5)`, `r round(confint(fit1)[2,2], 5)`)$, we can be reasonably confident that a positive correlation does exist between the number of reported UFO sightings in a state since 1974 and the mean population of that state since 1970.

We can plot this relationship using `ggplot`.

```{r}
ggplot(data=df, aes(x=pop, y=sightings)) + geom_point() + stat_smooth(method='lm', se=F)
```

A number of obvious outliers are noticeable, but otherwise the relationship appears linear.
We can graphically futher verify the assumptions of the least squares regression we performed.

```{r}
plot(fit1)
```

From the Q-Q plot, we can see that the errors do not appear to be normally distributed, with very large deviations from normal at either end.
Additionally, two major outliers with IDs 5 and 32 are visible on the plot of residuals vs leverage. We can identify which states have the largest prediction error.

```{r}
(top_resid <- fit1$residuals  %>%
    abs() %>%
    order(decreasing=T) %>%
    head(7) %>%
    df[., "name"])


ggplot(data=df, aes(x=pop, y=sightings)) +
    geom_point() +
    stat_smooth(method='lm', se=F) +
    geom_text_repel(aes(label=name), data=df[df$name %in% top_resid,])

```

From this we can clearly see California (point number 5 on the diagnostic plots) is a major outlier.
TODO: More analysis of this here


### Number of US Air Force bases vs. population

```{r}
(s2 <- summary(fit2 <- lm(base_count ~ pop, data=df)))
```

With an 95% confidence interval for $\beta_{pop} = (`r round(confint(fit2)[2,1], 10)`, `r round(confint(fit2)[2,2], 10)`)$, we can confidently state that there exists a positive relationship between the number of US military bases in a state and the mean population of that state from 1970-2010.

```{r}
ggplot(data=df, aes(x=pop, y=base_count)) + geom_point() + stat_smooth(method='lm', se=F)
```

From this graph we can see a loose but definite positive relationship between the two variables.

```{r}
plot(fit2)
```

There is a clear pattern in the plot of residuals vs fitted values due to the small number of discrete values which `base_count` can take on.
The normal Q-Q plot is almost a straight line, except at the high end where the residuals are larger than would be expected for a normal distribution.
In the plot of residuals vs leverage, California (point number 5) once again stands out as a significant outlier.
TODO: Analyze more

Additionally, it appears as though the variance in this relationship may be heteroscedastic.
We can test for this using the Breusch-Pagan test

```{r}
bptest(fit2)
```

Although the error appears heteroscedastic, according to the test results there is no significant evidence of heteroscedasticity.

Knowing that a relationship exists between `pop` and `sightings`, as well as `base_count` and `pop`, we can normalize the number of sightings by the population before fitting against `base_count` to account for these relationships. (TODO: can we?)

### Number of USAF bases per state vs sightings per thousand people per state

```{r}
### Normalize the number of sightings by the population of each state
df$sightings_per_thousand <- (df$sightings / df$pop) * 1000

(s3 <- summary(fit3 <- lm(sightings_per_thousand ~ base_count, data=df)))
```

With an 95% confidence interval for $\beta_{base\_count} = (`r round(confint(fit3)[2,1], 3)`, `r round(confint(fit3)[2,2], 3)`)$, we can confidently say there is no relationship between the number of US Air Force bases in a state and the population-adjusted number of reported UFO sightings in that state since 1970.

TODO: Analyze this relationship further

```{r}
ggplot(data=df, aes(x=base_count, y=sightings_per_thousand)) + geom_point() + stat_smooth(method='lm', se=F)
```

Graphically we can confirm the F-test results that there is little relationship between the two variables.


```{r}
plot(fit3)
```
