---
title: "Eda_Movies"
author: "Shane Caldwell"
date: "11/25/2017"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
#pull in alien movies
alien_movies <- read.csv(file = "/Users/shane/Documents/UFOTracker/data/raw/alien_movies_per_year.csv", header = TRUE)

#pull in sightings
sightings <- read.csv(file = "/Users/shane/Documents/UFOTracker/data/raw/ufo_sightings.csv", header = TRUE)

```

```{r}
ggplot(alien_movies, aes(x =year, y =num_movies )) + geom_point(shape=1) + labs( title = "Number of Alien Movies Per Year\n1902 - 2018", x = "Year", y = "Alien Movies Released") + theme(plot.title = element_text(hjust = 0.5))
```

It might be nice to normalize this data somehow. Maybe the percentage of alien movies out of all movies out? Right now we see a kind of linear relationship as the number of movies tends to increase over time. Maybe this will become less apparent if we can get rid of those zeros.

Mackenzie and I had discussed removing all data prior to 1974. This is because the UFO reporting society was founded in 1974, and they wouldn't have live records previous to that. It would all be "popular" events being recorded after the fact.


```{r}
alien_movies_new <- filter(alien_movies, year >= 1974)


ggplot(alien_movies_new, aes(x =year, y =num_movies )) + geom_point(shape=1) + labs( title = "Number of Alien Movies Per Year\n1974 - 2018", x = "Year", y = "Alien Movies Released") + theme(plot.title = element_text(hjust = 0.5))

```

Still a definite trend upwards. That could have to do with the number of movies released increasing in general moreso than an increase in the genre alien movies specifically. 

Anyway, we have to filter the sightings dataset to find more specific information about our dataset. Currently we can't compare the number of movies to our number of sightings. 

```{r}

fl_sightings <- filter(sightings, tolower(state) == 'fl' )
fl_sightings$date_time <- as.Date(fl_sightings$date_time, "%m/%d/%y")
tmp <- lapply(strsplit(as.character(fl_sightings$date_time), "-"), `[[`, 1)
tmp2 <- sapply(tmp, "[[", 1)
fl_sightings$year <- as.numeric(tmp2)
fl_sightings_75 <- filter(fl_sightings, year >= 1975)
fl_sightings_75 <- filter(fl_sightings, year < 2018)

sightings_per_year_fl <- as.data.frame(table(fl_sightings_75$year))

#sightings_per_year_fl_75$Freq
df1 <- cbind(sightings_per_year_fl$Freq[5:49], alien_movies_new$num_movies)
df1 <- as.data.frame(df1)

colnames(df1) <- c("Sightings_Per_Year", "Alien_Movies_Per_Year")

ggplot(df1, aes(x =Alien_Movies_Per_Year, y =Sightings_Per_Year )) + geom_point(shape=1) + labs( title = "Number of Alien Sightings VS Alien Movies eleased\n1974 - 2018", x = "Alien Movies Released Per Year", y = "UFO Sightings Per Year") + theme(plot.title = element_text(hjust = 0.5))

```

Nothing that seems like a line could be drawn through it.  


```{r}
lm.movies <- lm(Sightings_Per_Year~Alien_Movies_Per_Year, data = df1)
summary(lm.movies)
```
Results here are interesting. Our R-squared isn't incredible, only capturing 19% of the variance in our data. Without other variables in the model, it's difficult to say if the alien movies per year are useful on their own. 

### Model Diagnostics

```{r}
library(olsrr)
ols_bp_test(lm.movies)
```
Here we do not reject the probability that the variance is constant. But is this test really capturing the results of the data?

```{r}
residuals <- resid(lm.movies)
plot(residuals)
```

This definitely does not appear to have no pattern! We could easily draw a curve fitting this data. The variance increases the further we go through the function.

```{r}
qqnorm(lm.movies$residuals)
qqline(lm.movies$residuals)
```

The qqplot here suggests our data is not normally distributed. As you can see, the forced normalized values and the given values don't match up. The qqplot indicates significant deviation from normality.

## Test for autocorrelation

```{r}
library(lmtest)
results = dwtest(Sightings_Per_Year~Alien_Movies_Per_Year, data = df1)
```

There definitely appears to be autocorrelation in the residuals. 

```{r}
myts <- ts(df1$Sightings_Per_Year, start=c(1975, 1), end=c(2017), frequency=1)
plot(myts)
```

## Dickey Fuller Test to determine if data is stationary 

```{r}
library(tseries)
adf.test(myts, k = 3)
```
```{r}

```

