---
title: "EDA Beer & Alien Sightings"
author: "Fanny Chow"
output:
  pdf_document: default
  html_document:
      keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(tidyr)
```



```{r}
# set paths to data source & read in files

setwd("~/Google Drive/stat/UFOTracker")
my.path <- "~/Google Drive/stat/UFOTracker"
beer.path <- "data/raw/brew_count_by_state_1984_2017.csv"
sightings.path <- "data/raw/ufo_sightings.csv"

beer.raw <- fread(file.path(my.path, beer.path), header=TRUE, na.strings=c("*", ""))
sightings.raw <- fread(file.path(my.path, sightings.path), header = TRUE, na.strings = c("", "Unknown", "--"))

```


```{r}
# clean up junk at bottom file
beer.db <- beer.raw %>% 
  filter(!is.na(STATE)) %>%
  filter(STATE != "Total") %>% 
  filter(STATE != "Other") %>% 
  filter(STATE != "* No reportable data") %>% 
  filter(STATE != "«This list will be updated quarterly.")
```

```{r}
# create proper data types
beer.db$STATE <- as.factor(beer.db$STATE)

# wide to long format
olddata_wide <- beer.db
keycol <- "year"
valuecol <- "breweries"
gathercols <- as.character(seq(1984, 2017))

beer.df <- gather_(olddata_wide, keycol, valuecol, gathercols)

```

At a high-level glance, the general trend is increase number of breweries through the years for each s tate. Note that the number of breweries in 2005 will be contingent on the number of breweries in 2004, and there will be autocorrolation through years.   
```{r}
beer.df$breweries <- as.numeric(beer.df$breweries)
```

```{r}
#breweries.year <- ggplot(beer.df, aes(x = year, y = breweries, group=1))
#breweries.year + geom_point() + geom_line(aes(color = STATE))  

ggplot() +
  geom_point(data=beer.df, aes(year, breweries, color=STATE))
```


Let's take a look at the number of breweries in each state through the years sorted by states. Since there's over 50 states we're looking at, it's challenging to discern trends from looking at all the states at once.  
```{r}
breweries.year <- ggplot(beer.df, aes(x = year, y = breweries, group=1))
(p2 <- breweries.year + geom_line() +
   facet_wrap(~STATE, ncol = 10))
```


Let's focus on the state of Florida through the years. We observe an upward trend and then a sudden dip from the late 90's to 2010.  
```{r}
florida.brew.df <- beer.df %>% 
  filter(STATE == "FL")
florida.brew.df

plot(florida.brew.df$year, florida.brew.df$breweries)

```








Since the range of time in the breweries data is from 1984-2017, let's subset the equivalent years from the sightings data.  
```{r}
# clean up sightings data
fl.sightings <- sightings.raw %>% 
  filter(state == 'FL') %>% 
  mutate(year = as.numeric(format(as.Date(date_time, format="%m/%d/%y"),"%Y"))) %>% 
  filter(year >= 1984) %>% 
  filter(year <= 2017) %>% 
  #group_by(year) %>% 
  count(year) %>% 
  rename(sightings_year = n)
 # mutate(sightings_year = n())
#group_by(`Student ID`) %>%
 # mutate(`Dupe Check`= n())

fl.sightings
#fl.sightings$year <- as.numeric(format(as.Date(fl.sightings$date_time, format="%m/%d/%y"),"%Y"))
#filter(fl.sightings, year > 1983)


#fl.sightings.count <- as.data.frame(table(fl.sightings$year))
#colnames(fl.sightings.count) <- c("year","sightings_per_year")
#as.numeric(fl.sightings.count$year)
#str(fl.sightings.count)

```

Let's take a snapshot of sightings per year in Florida.  
```{r}
plot(fl.sightings$year, fl.sightings$sightings_year, type="l")
```



Let's compare the 2 plots at once. Interesting how the 2 plots follow the same shape until the early 2000s and then diverge drastically afterwards. It makes sense to see an overall trend of increased interest in breweries over time.   
```{r}
plot(florida.brew.df$year, florida.brew.df$breweries, type = "l", col="blue")
par(new=TRUE)
plot(fl.sightings$year, fl.sightings$sightings_year, type="l", col="orange")
legend("topleft",legend=c("Breweries", "Sightings"),
       col=c("blue", "orange"), lty=1:2, cex=0.8)

```


In order to stabilize the variance, we will transform the data with a log transformation. Notice how the the residuals do not vary as much and the variance is closer to constant now.  

```{r}

```





Let's use check for autocorrelation between the years using a statistical test. 






